---
title: "MLH05"
author: "Liu Zhe"
date: "2022-11-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = TRUE,
	warning = FALSE
)
```

# Environment Variables and Packages

```{r}
options(java.parameters = "-Xmx2048m",
        stringsAsFactors = FALSE, 
        encoding = 'UTF-8')

suppressPackageStartupMessages({
  # ISLR2
  library(ISLR2)
  # Splines
  library(splines)
  # DM
  library(zip)
  library(openxlsx)
  library(readxl)
  library(writexl)
  library(RcppRoll)
  library(plyr)
  library(stringi)
  library(feather)
  library(RODBC)
  library(MASS)
  library(car)
  library(caret)
  library(data.table)
  library(lubridate)
  library(plotly)
  library(pROC)
  library(tidymodels)
  library(tidyverse)
})
```

# Boston

Select particular variables of Boston data. 

```{r}
boston.subset <- Boston %>% 
  select(dis, nox) %>% 
  arrange(dis)

head(boston.subset)
```

### Cubic polynomial regression. 

Fit. 

```{r}
boston.poly <- lm(nox ~ poly(dis, degree = 3), data = boston.subset)
summary(boston.poly)
```

Fitted values. 

```{r}
boston.poly.pred <- boston.subset %>% 
  mutate(nox_pred = boston.poly$fitted.values)
```

Fit plot. 

```{r}
boston.poly.plot <- plot_ly(data = boston.poly.pred, x = ~dis) %>% 
  add_trace(y = ~nox, name = 'True', type = 'scatter', mode = 'markers') %>% 
  add_trace(y = ~nox_pred, name = 'Fit', type = 'scatter', mode = 'lines') %>% 
  layout(
    title = 'Polynomial Fits'
  )
boston.poly.plot
```

### Polynomial fits of different degrees

Fit. 

```{r}
boston.poly.list <- list()
for (i in 1:10) {
  boston.poly.list[[i]] <- lm(nox ~ poly(dis, degree = i), data = boston.subset)
}
```

Fitted values. 

```{r}
boston.poly.list.pred <- boston.subset
for (i in 1:10) {
  boston.poly.list.pred[paste0('nox_pred', str_pad(i, 2, side = 'left', pad = 0))] <- boston.poly.list[[i]]$fitted.values
}
```

Fit plot. 

```{r}
pred.name <- paste0('nox_pred', str_pad(1:10, 2, side = 'left', pad = 0))

boston.poly.list.plot <- plot_ly(data = boston.poly.list.pred, x = ~dis) %>% 
  add_trace(y = ~nox, name = 'nox', type = 'scatter', mode = 'markers') %>% 
    add_trace(y = ~get(pred.name[1]), name = pred.name[i], 
              type = 'scatter', mode = 'lines') %>% 
    add_trace(y = ~get(pred.name[2]), name = pred.name[i], 
              type = 'scatter', mode = 'lines') %>% 
    add_trace(y = ~get(pred.name[3]), name = pred.name[i], 
              type = 'scatter', mode = 'lines') %>% 
    add_trace(y = ~get(pred.name[4]), name = pred.name[i], 
              type = 'scatter', mode = 'lines') %>% 
    add_trace(y = ~get(pred.name[5]), name = pred.name[i], 
              type = 'scatter', mode = 'lines') %>% 
    add_trace(y = ~get(pred.name[6]), name = pred.name[i], 
              type = 'scatter', mode = 'lines') %>% 
    add_trace(y = ~get(pred.name[7]), name = pred.name[i], 
              type = 'scatter', mode = 'lines') %>% 
    add_trace(y = ~get(pred.name[8]), name = pred.name[i], 
              type = 'scatter', mode = 'lines') %>% 
    add_trace(y = ~get(pred.name[9]), name = pred.name[i], 
              type = 'scatter', mode = 'lines') %>% 
    add_trace(y = ~get(pred.name[10]), name = pred.name[i], 
              type = 'scatter', mode = 'lines') %>% 
  layout(
    title = 'Polynomial Fits of Different Degrees'
  )
boston.poly.list.plot
```

Residual sum of squares. 

```{r}
boston.poly.list.res <- boston.subset
for (i in 1:10) {
  boston.poly.list.res[paste0('nox_res', str_pad(i, 2, side = 'left', pad = 0))] <- boston.poly.list[[i]]$residuals
}

boston.poly.list.res2 <- boston.poly.list.res %>% 
  select(starts_with('nox_')) %>% 
  summarise_all(function(x) {sum(x^2)}) %>% 
  as.data.table() %>% 
  melt(variable.name = 'poly', value.name = 'rss')
boston.poly.list.res2
```

### Cross-validation

Split data set to 10-fold. 

```{r}
set.seed(0803)
folds.index <- createFolds(1:nrow(boston.subset), k = 10)
```

CV. 

```{r}
rss.degrees <- c()
for (i in 1:10) {
  rss.folds <- c()
  
  for (j in 1:10) {
    boston.train <- boston.subset %>% 
      filter(!(row_number() %in% folds.index[[j]]))
    
    boston.valid <- boston.subset %>% 
      filter(row_number() %in% folds.index[[j]])
    
    boston.poly.cv <- lm(nox ~ poly(dis, degree = i), data = boston.train)
    
    rss <- sum((boston.valid$nox - predict(boston.poly.cv, newdata = boston.valid['dis'])) ^ 2)
    rss.folds <- append(rss.folds, rss)
  }
  
  rss.degrees <- append(rss.degrees, sum(rss.folds))
}

rss.degrees
```

When degree = 3, the polynomial regression has the minimal RSS, so the optimal degree is 3. 

### Regression spline

```{r}
boston.sp <- lm(nox ~ bs(dis, df = 4, knots = 2, degree = 3), data = boston.subset)
```










